The Bose-Hubbard model simulates identical, interacting bosons on lattice. These bosons (particles with whole integer spin) are allowed to live in the same quantum state as other bosons. In the case of the system we looked into, this meant that the bosons were allowed to live on the same site as other bosons. Therefore, an infinite number of bosons could live on the same site at the same time. This system of bosons on a lattice has applications in quantum computation through its use in describing Rydberg atom quantum hardware, where Rydberg atoms are a specific subset of bosons (atoms in this case). To quote Wu in his 2021 paper, "More generally, any atom (moleculeor semiconductor quantum dot) in a state with a highly-excited electron, i.e., with valance electron in a large principal quantum number \textit{n} state of several tens to hundreds or even higher, is regarded as Rydberg atom."\cite{wu_concise_2021}

The following subsection consists of the by-hand exact-diagonalization calculations I preformed during my research experience at the University of Tennessee Knoxville in the summer of 2023. These calculations are split into two main pieces: the first is for first quantization and the second for second quantization. First quantization consists of labeling the system according to the sites where the bosons exist. Second quantization consists of labeling the system according to the individual particles themselves.

\subsection{Bose-Hubbard Calculations} \label{calculations}
\subsubsection{$2^{nd}$ Quantization Von Nuemann Entanglement Entropy Calculation}
To get a better initial feeling for this calculation, we begin with the simpler case: $2^{nd}$ quantization or spatial entanglement. Additionally, systems bigger than 2 particles on 2 sites make by-hand calculations unreadable. Therefore, all following calculations and data analysis only deal with 2 particles on 2 sites for the Bose-Hubbard model. The basis for $2^{nd}$ quantization with 2 particles on 2 sites is as follows:
\begin{equation}
|2 0 \rangle, |1 1 \rangle, |0 2 \rangle,
\end{equation}
\noindent where the left-most number corresponds to the number of bosons on site A and the right-most on site B. The full Hamiltonian for the Bose-Hubbard system is
\begin{equation}
\hat{H} = -J\sum_i{b_i^{\dagger}b_{i+1} + b_{i+1}^{\dagger}b_i} + \frac{U}{2}\sum_i{n_i(n_i - 1)} - \mu\sum_i{n_i},
\label{eq:2}
\end{equation}
\noindent where J is the hopping term, U is the potential energy, and $\mu$ is the chemical potential. $b^{\dagger}$ and $b$ are the creation and annihilation operators, respectively. They act as follows:
\begin{align}
b^{\dagger}|n\rangle &= \sqrt{n+1}\enspace|n+1\rangle \\
b|n\rangle &= \sqrt{n}\enspace|n-1\rangle.
\end{align}
\noindent J, the hopping term, is also denoted as the tunneling term. This term governs how well the particles are able to move between sites. U, the potential energy term, governs the repulsive force between the bosons. This repulsive potential is a result of the bosons approaching each other in close proximity to where their electrons clouds interact, forcing the bosons in opposite directions. 
\noindent The full Hamiltonian can be calculated by calculating the kinetic and potential pieces separately, then simply adding them together. Start by calculating the kinetic energy piece of the full matrix Hamiltonian from \cref{eq:2}. The kinetic energy part of the full hamiltonian above for the 2 particle, 2 site system is
\begin{equation}
\hat{H}_{KE}=-J(b_1^{\dagger}b_2+b_2^{\dagger}b_1).
\end{equation}
\noindent The matrix elements of the kinetic energy part of the Hamiltonian are calculated as follows:
\begin{align*}
\langle{20|H_{KE}|20\rangle} &=0 \\
\langle{11|H_{KE}|20\rangle} &=-\sqrt{2}J \\
\langle{02|H_{KE}|20\rangle} &=0 \\
\langle{20|H_{KE}|11\rangle} &=-\sqrt{2}J \\
\langle{11|H_{KE}|11\rangle} &=0 \\
\langle{02|H_{KE}|11\rangle} &=-\sqrt{2}J \\
\langle{20|H_{KE}|02\rangle} &=0 \\
\langle{11|H_{KE}|02\rangle} &=-\sqrt{2}J \\
\langle{02|H_{KE}|02\rangle} &=0.
\end{align*}
\noindent From these matrix elements, the full kinetic energy piece of the Hamiltonian is written as:
\begin{equation}
\hat{H}_{KE} = \begin{pmatrix} 0 & -\sqrt{2} & 0 \\ -\sqrt{2} & 0 & -\sqrt{2} \\ 0 & -\sqrt{2} & 0 \end{pmatrix}.
\end{equation}
\noindent Next, the potential part of the full Hamiltonian from \cref{eq:2} can be calculated as follows:
\begin{equation}
\hat{H}_{P} = \frac{U}{2}\sum_i{n_i(n_i - 1)} = \frac{U}{2}[n_1(n_1 - 1) + n_2(n_2 - 1)],
\end{equation}
\noindent where $n_i = b_i^{\dagger}b_i$. The elements of the potential part of the Hamiltonian matrix are:
\begin{align*}
\langle{20|H_{P}|20\rangle} &=U \\
\langle{11|H_{P}|20\rangle} &=0 \\
\langle{02|H_{P}|20\rangle} &=0 \\
\langle{20|H_{P}|11\rangle} &=0 \\
\langle{11|H_{P}|11\rangle} &=0 \\
\langle{02|H_{P}|11\rangle} &=0 \\
\langle{20|H_{P}|02\rangle} &=0 \\
\langle{11|H_{P}|02\rangle} &=0 \\
\langle{02|H_{P}|02\rangle} &=U.
\end{align*}
\noindent From these matrix elements, the full potential energy part of the Hamiltonian is written as:
\begin{equation}
\hat{H}_{P} = \begin{pmatrix} U & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & U \end{pmatrix}.
\end{equation}
\noindent The separate kinetic and potential energy pieces of the matrix Hamiltonian can be combined together to make the full matrix as follows:
\begin{equation}
\hat{H} = \begin{pmatrix} U & -\sqrt{2} & 0 \\ -\sqrt{2} & 0 & -\sqrt{2} \\ 0 & -\sqrt{2} & U \end{pmatrix}.
\end{equation}
\noindent Now that we have the full matrix Hamiltonian, we can find the energy states of the system by solving the Schr\"{o}dinger equation. The general form of the Schr\"{o}dinger equation is
\begin{equation}
\hat{H}|\psi\rangle = E|\psi\rangle.
\end{equation}
\noindent Where $E$ is the energy of the system and $|\psi\rangle$ is the wavefunction that describes the system.
\noindent This equation can then be manipulated to look like the following equation:
\begin{equation}
\left( \hat{H}-E\hat{I} \right) |\psi\rangle = 0.
\end{equation}
\noindent Since we are not interested in the solution where $|\psi\rangle=0$, then $\hat{H}-E\hat{I}$ must be zero. This yields the following characteristic equation:
\begin{equation}
\hat{H}-E\hat{I} = 0.
\end{equation}
\noindent This characteristic equation can be solved via expanding the matrix minors and solving the resulting polynomial, where
\begin{equation}
-E^2 + 2UE^2 + 4J^2E - U^2E - 4J^2U = 0.
\label{eq:13}
\end{equation}
\noindent Solving this equation yields 
\begin{align*}
E_1 &= U \\
E_2 &= \frac{1}{2} \left( U - \sqrt{16J^2 + U^2} \right) \\
E_3 &= \frac{1}{2} \left( U + \sqrt{16J^2 + U^2} \right).
\end{align*}
\noindent The solutions to \cref{eq:13} are listed above and are the eigenvalues for the Hamiltonian.  Looking at the three energies listed above, it is obvious that $E_2$ is the ground state energy since it is the lowest of the three. Before the entanglement entropy can be calculated, the density matrix and reduced density matrix must be found. The full density matrix is defined as $\rho = |\psi \rangle \langle \psi|$. The calculation is as follows:
\begin{align}
\hat{\rho} = \left( \alpha |2 0 \rangle + \beta |1 1 \rangle + \gamma |0 2 \rangle \right) \left( \alpha^* \langle 2 0 | + \beta^* \langle 1 1| + \gamma^* \langle 0 2| \right).
\end{align}
\noindent In matrix notation:
\begin{equation}
\hat{\rho} = \begin{bmatrix} \alpha \alpha^* & \alpha \beta^* & \alpha \gamma^* \\ \beta \alpha^* & \beta \beta^* & \beta \gamma^* \\ \gamma \alpha^* & \gamma \beta^* & \gamma \gamma^* \end{bmatrix}.
\end{equation}
\noindent The reduced density matrix is calculated as follows:
\begin{align}
\hat{\rho}_A &= \sum_{n=0}^{2}{{}_B \langle n | \psi \rangle \langle \psi | n \rangle_B} = {}_B \langle 0 | \psi \rangle \langle \psi | 0 \rangle_B + {}_B \langle 1 | \psi \rangle \langle \psi | 1 \rangle_B + {}_B \langle 2 | \psi \rangle \langle \psi | 2 \rangle_B \\
\hat{\rho}_A &=|\alpha|^2 {}_A |2\rangle\langle2|_A + |\beta|^2 {}_A |1\rangle\langle1|_A + |\gamma|^2 {}_A |0\rangle\langle0|_A.
\end{align}
\noindent These elements written in matrix notation are given by:
\begin{equation}
\hat{\rho}_A = \begin{bmatrix} |\alpha|^2 & 0 & 0 \\ 0 & |\beta|^2 & 0 \\ 0 & 0 & |\gamma|^2 \end{bmatrix}
\end{equation}
\noindent Now, we want to calculate the ground state eigenvector. This can be done by substituting our ground state eigenvalue into equation [10] and solving the resulting augmented matrix. This matrix is as follows:
\begin{equation}
\begin{bmatrix} U - \frac{1}{2} \left( U - \sqrt{16J^2 + U^2} \right) & -\sqrt{2}J & 0 \\ -\sqrt{2}J & -\frac{1}{2} \left( U - \sqrt{16J^2 + U^2} \right) & -\sqrt{2}J \\ 0 & -\sqrt{2}J & U - \frac{1}{2} \left( U - \sqrt{16J^2 + U^2} \right) \end{bmatrix} \begin{bmatrix} \alpha \\ \beta \\ \gamma \end{bmatrix} = 0.
\end{equation}
\noindent Solving the above equation provides the solutions for $\alpha$, $\beta$, \& $\gamma$ as follows:
\begin{align*}
\alpha &= \frac{2}{\sqrt{ U'^2 + U' \sqrt{U' + 16} + 16}} = \gamma \\
\beta &= \frac{U' + \sqrt{U' + 16}}{\sqrt{2}\sqrt{U'^2 + U' \sqrt{U' + 16} + 16}},
\end{align*}
\noindent where $U' = \frac{U}{J}$. Finally, the Von Nuemann spatial entanglement entropy is calculated as follows:
\begin{align}
S_1 &= -\text{Tr} \left( \rho_A \ln{\rho_A} \right) = -\text{Tr} \begin{bmatrix} |\alpha|^2 \ln{|\alpha|^2} & 0 & 0 \\ 0 & |\beta|^2 \ln{|\beta|^2} & 0 \\ 0 & 0 & |\gamma|^2 \ln{|\gamma|^2} \end{bmatrix} \\
S_1 &= -\left[ |\alpha|^2 \ln{|\alpha|^2} + |\beta|^2 \ln{|\beta|^2} + |\gamma|^2 \ln{|\gamma|^2} \right]
\end{align}
\noindent This result shows that the Von Neumann entanglement entropy depends only on the elements of the ground state vector of the system: $\alpha$, $\beta$, and $\gamma$.
\subsubsection{$1^{st}$ Quantization Von Nuemann Entanglement Entropy Calculation}
Having obtained the Von Neumann entanglement entropy for $2^{st}$ quantization, we want to run through the same process with $1^{st}$ quantization to obtain the particle entanglement. In the case of $1^{st}$ quantization, we label the particles rather than the sites. The basis for $1^{st}$ quantization is as follows:
\begin{align}
|1_1 2_1 \rangle, |1_1 2_2 \rangle, |1_2 2_1 \rangle, |1_2 2_2 \rangle,
\end{align}
\noindent where the state numbers (i.e., the base numbers) are the particle labels and the subscripts are the site labels (i.e., which site the particle exists on). We begin by calculating the full Hamiltonian, which behaves the same way for $1^{st}$ quantization as it did for $2^{nd}$ quantization:
\begin{equation}
\hat{H} = \begin{bmatrix} \langle 1_1 2_1 | \hat{H} | 1_1 2_1 \rangle & \langle 1_1 2_1 | \hat{H} | 1_1 2_2 \rangle & \langle 1_1 2_1 | \hat{H} | 1_2 2_1 \rangle & \langle 1_1 2_1 | \hat{H} | 1_2 2_2 \rangle \\ \langle 1_1 2_2 | \hat{H} | 1_1 2_1 \rangle & \langle 1_1 2_2 | \hat{H} | 1_1 2_2 \rangle & \langle 1_1 2_2 | \hat{H} | 1_2 2_1 \rangle & \langle 1_1 2_2 | \hat{H} | 1_2 2_2 \rangle \\ \langle 1_2 2_1 | \hat{H} | 1_1 2_1 \rangle & \langle 1_2 2_1 | \hat{H} | 1_1 2_2 \rangle & \langle 1_2 2_1 | \hat{H} | 1_2 2_1 \rangle & \langle 1_2 2_1 | \hat{H} | 1_2 2_2 \rangle \\ \langle 1_2 2_2 | \hat{H} | 1_1 2_1 \rangle & \langle 1_2 2_2 | \hat{H} | 1_1 2_2 \rangle & \langle 1_2 2_2 | \hat{H} | 1_2 2_1 \rangle & \langle 1_2 2_2 | \hat{H} | 1_2 2_2 \rangle \end{bmatrix} 
\end{equation}
\begin{equation}
\hat{H} = \begin{bmatrix} U & -\sqrt{2} & -\sqrt{2} & 0 \\ -\sqrt{2} & 0 & 0 & -\sqrt{2} \\ -\sqrt{2} & 0 & 0 & -\sqrt{2} \\ 0 & -\sqrt{2} & -\sqrt{2} & U \end{bmatrix}.
\end{equation}
\noindent The full density matrix is then given by $|\psi \rangle \langle \psi|$, where $\psi$, the ground state, is given by:
\begin{equation*}
\psi = \alpha |1_1 2_1 \rangle + \frac{\beta}{\sqrt{2}} \left( |1_1 2_2 \rangle + |1_2 2_1 \rangle \right) + \gamma |1_2 2_2 \rangle,
\end{equation*}
\iffalse
\begin{equation}
\hat{\rho} = \begin{bmatrix} \langle 1_1 2_1 | \hat{\rho} | 1_1 2_1 \rangle & \langle 1_1 2_1 | \hat{\rho} | 1_1 2_2 \rangle & \langle 1_1 2_1 | \hat{\rho} | 1_2 2_1 \rangle & \langle 1_1 2_1 | \hat{\rho} | 1_2 2_2 \rangle \\ \langle 1_1 2_2 | \hat{\rho} | 1_1 2_1 \rangle & \langle 1_1 2_2 | \hat{\rho} | 1_1 2_2 \rangle & \langle 1_1 2_2 | \hat{\rho} | 1_2 2_1 \rangle & \langle 1_1 2_2 | \hat{\rho} | 1_2 2_2 \rangle \\ \langle 1_2 2_1 | \hat{\rho} | 1_1 2_1 \rangle & \langle 1_2 2_1 | \hat{\rho} | 1_1 2_2 \rangle & \langle 1_2 2_1 | \hat{\rho} | 1_2 2_1 \rangle & \langle 1_2 2_1 | \hat{\rho} | 1_2 2_2 \rangle \\ \langle 1_2 2_2 | \hat{\rho} | 1_1 2_1 \rangle & \langle 1_2 2_2 | \hat{\rho} | 1_1 2_2 \rangle & \langle 1_2 2_2 | \hat{\rho} | 1_2 2_1 \rangle & \langle 1_2 2_2 | \hat{\rho} | 1_2 2_2 \rangle \end{bmatrix} 
\end{equation}
\fi
\noindent where $\alpha$, $\beta$, and $\gamma$ are the elements of the ground state vector of the system. The full density matrix is
\begin{equation}
\hat{\rho} = \begin{bmatrix} |\alpha|^2 & \frac{\alpha \beta^*}{\sqrt{2}} & \frac{\alpha \beta^*}{\sqrt{2}} & \alpha \gamma^* \\ \frac{\beta \alpha^*}{\sqrt{2}} & \frac{|\beta|^2}{2} & \frac{|\beta|^2}{2} & \frac{\beta \gamma^*}{\sqrt{2}} \\ \frac{\beta \alpha^*}{\sqrt{2}} & \frac{|\beta|^2}{2} & \frac{|\beta|^2}{2} & \frac{\beta \gamma^*}{\sqrt{2}} \\ \gamma \alpha^* & \frac{\gamma \beta^*}{\sqrt{2}} & \frac{\gamma \beta^*}{\sqrt{2}} & |\gamma|^2 \end{bmatrix}.
\end{equation}
\noindent We then needed to calculate the reduced density matrix, which is given by tracing out the degrees of freedom of the second particle. The reduced density matrix is as follows:
\begin{align}
\hat{\rho}_A &= \sum_{n=1}^{2} {}_B \langle n|\psi \rangle \langle \psi|n \rangle_B = {}_B\langle 2_1|\psi \rangle \langle \psi|2_1 \rangle_B + {}_B\langle 2_2|\psi \rangle \langle \psi|2_2 \rangle_B.
\end{align}
The only surviving terms from the above calculation are
\begin{align*}
\hat{\rho}_A &= \left( |\alpha|^2 + \frac{|\beta|^2}{2} \right) {}_A |1_1 \rangle \langle 1_1|_A
+ \left( \frac{\alpha \beta^* + \beta \gamma^*}{\sqrt{2}} \right) {}_A |1_1 \rangle \langle 1_2|_A \\
&+ \left( \frac{\beta \alpha^*+ \gamma \beta^*}{\sqrt{2}} \right) {}_A |1_2 \rangle \langle 1_1|
+ \left( \frac{|\beta|^2}{2} + |\gamma|^2 \right) {}_A |1_2 \rangle \langle 1_2|_A.
\end{align*}
\begin{equation}
\hat{\rho}_A = \begin{bmatrix} {}_A \langle 1_1 | \hat{\rho_A} | 1_1 \rangle & {}_A \langle 1_1 | \hat{\rho_A} | 1_2 \rangle \\ {}_A \langle 1_2 | \hat{\rho_A} | 1_1 \rangle & {}_A \langle 1_2 | \hat{\rho_A} | 1_2 \rangle \end{bmatrix} = \begin{bmatrix} |\alpha|^2 + \frac{|\beta|^2}{2} & \frac{\alpha \beta^* + \beta \gamma^*}{\sqrt{2}} \\ \frac{\beta \alpha^* + \gamma \beta^*}{\sqrt{2}} & \frac{|\beta|^2}{2} + |\gamma|^2 \end{bmatrix} \\
\hat{\rho}_A = \begin{bmatrix} \frac{1}{2} & \sqrt{2} \alpha \beta \\ \sqrt{2} \alpha \beta & \frac{1}{2} \end{bmatrix}.
\end{equation}
\noindent To get the Von Nuemann entanglement entropy, the eigenvalues of the above matrix are needed. To obtain these eigenvalues, the following equation must be solved:
\begin{equation}
\hat{\rho}_A - \lambda \hat{I} = 0, \\
\end{equation}
\noindent where $\lambda$ is the eigenvalue we are solving for. In matrix form, this becomes:
\begin{equation}
\begin{bmatrix} \frac{1}{2} - \lambda & \sqrt{2} \alpha \beta \\ \sqrt{2} \alpha \beta & \frac{1}{2} - \lambda \end{bmatrix} = 0.
\end{equation}
\noindent Taking the determinant of this equation yields the characteristic equation:
\begin{align}
\lambda^2 - \lambda - 2 \alpha^2 + 4 \alpha^4 + \frac{1}{4} &= 0.
\end{align}
\noindent The eigenvalues can then be calculated using the quadratic equation:
\begin{align}
\lambda_{\pm} &= \frac{1 \pm \sqrt{1^2 - 4 \left( 1 \right) \left( -2 \alpha^2 + 4 \alpha^4 + \frac{1}{4} \right) }}{2} = \frac{1 \pm 2 \alpha \sqrt{2 - 4 \alpha^2}}{2}.
\label{eq:lambda_pm}
\end{align}
\noindent The Von Nuemann particle entanglement entropy is then calculated as follows:
\begin{align}
S_1 &= -\text{Tr} \left( \rho_A \ln{\rho_A} \right) = -\text{Tr} \begin{bmatrix} \lambda_+ & 0 \\ 0 & \lambda_- \end{bmatrix} \begin{bmatrix} \ln{\lambda_+} & 0 \\ 0 & \ln{\lambda_-} \end{bmatrix} \\
S_1 &= -\lambda_+ \ln{\lambda_+} - \lambda_- \ln{\lambda_-}
\end{align}
\noindent where $\lambda_+$ and $\lambda_-$ are defined by \cref*{eq:lambda_pm}. Similar to the Von Neumann entanglement entropy calculation for $2^{nd}$ quantization, the entanglement entropy for $1^{st}$ quantization also depends only on the elements of the ground state vector for the system.